---
title: "esm206_lab6"
author: "Shayan Kaveh"
date: "11/2/2021"
output: html_document
---

## Set Up

```{r setup, include = TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Rank based test example

Mann-Whitney U (non-parametric alternative to an unpaired 2-sample t-test)

H0: Equal ranks (or medians) HA: Ranks are not equal (or, medians are not equal)

### Create samples

`set.seed()` to create "pseudorandom, but reproducible samples":

```{r}
set.seed(1414) # this will create the same random sample for anyone who uses this same seed
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

Take a quick look at the distributions using `hist()`:

```{r}
hist(gp_1)
hist(gp_2)
```

Try a t-test to compare results:

```{r}
ttest <- 
  t.test(gp_1, gp_2) 
# if I wanted a paired t-test, I would need to include `paired = "TRUE"` within the t.test() function
```

This p-value means that IF our 2 samples are drawn from populations with the same mean (i.e. if the null hypothesis is true), there is `r round(ttest$p.value,3)*100`% probability of finding sample means *at least as different* as mine by random chance.


